---
title: 'Mixed Linear Model'
author: "Sungkyung Kang"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Install & load packages
```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(lme4)
library(lmerTest)
library(lsmeans)
library(nlme)
library(MuMIn)
library(dplyr)
library(multcomp)
library(arm)
library(xtable)
library(plyr)
```
Encoding: UTF-8

```{r}
rats$treatm <- factor(rats$treatm)
rats$cage <- factor(rats$cage)
rats$month <- factor(rats$month)
```
In the variance structure two observations from different cages are uncorrelated; 
two observations from the same cage are positively correlated. 

```{r}
mod0 <- lmerTest::lmer(lnc~month+treatm+month:treatm+(1|cage),data=rats)
arm::display(mod0) #summary(mod0)
```
`mod0` is a linear mixed model with an assumption of an (random) intercept is different for each **cage**. 
It coincides with the above formula. `(1|cage)` means `lmer` fits a linear model with a varying intercept group effect using the variable **cage**. `display` is a brief version of `summary`. From `mod0` I have tried to vary a slope in a model like `(1+treatm|cage)` or `(1+month|cage)`. These were in trouble due to scale problems. 

```{r}
mod1 <- lmerTest::lmer(lnc~month+treatm+month:treatm+(1|cage)+(1|month),data=rats)
arm::display(mod1)
```
When `mod0` is included a random effect `(1|treatm)` instead of `(1|month)`, it had an error. 
The model, `mod1` is to vary intercepts with two random effects **cage** and **month**.

```{r}
anova(mod0,mod1,refit=F) #mod0
```

```{r,results='asis'}
print(xtable(anova(mod0)),type="html")
```
The two models have same fixed effects but different random effects. 
In this case, if the models are fit by restricted maximum likelihood (REML) a likelihood ratio test is appropriate. 
The option `refit=F` is added in ANOVA. Of these two models, `mod0` is better than `mod1`. 

```{r}
mod2 <- nlme::gls(lnc~month+treatm+month:treatm,correlation=corCompSymm(form=~1|cage),data=rats)
mod2
intervals(mod2)
```
For specifying the correlation, `corCompSymm` is used as an option to `gls` function. `mod2` structure is same as `mod0` 
which is selected as a good model. 

###b.
Determine if change in activity over time depends on treatment.
What is the estimated change in activity per month for each treatment?
```{r}
mod3 <- lmerTest::lmer(lnc~month*treatm+(1|treatm/month),data=rats)
mod3
```

```{r,results='asis',warning=FALSE}
print(xtable(anova(mod3)),type="html")
```
`mod3` is assumed to use the nested data which **month** is nested by **treatm**. `(1|treatm/month)` 
means a mixed effect term for varying intercepts 1|by **treatm** and each **month** is nested within **treatm**. 
In this model the estimated change is the terms, **month** (9 dummies), **treatm** (2 dummies) and **month:treatm** (18 dummies).

```{r,results='asis',warning=FALSE}
print(xtable(anova(mod3)),type="html")
```

##Q2.Oats data
```{r}
Oats$nitro <- factor(Oats$nitro)
```

```{r}
ggplot(Oats,aes(x=yield,y=nitro,group=Variety))+
      geom_line(aes(col=Block))+
      facet_grid(Variety~Block,scales="free")+
      labs(title="The Split Plot of Oats")+
      guides(col=F)
```
<br><br>
The range of **yield** (x-axis) is depended on the levels of **Variety** and **Block**. 

```{r}
ggplot(Oats,aes(x=nitro,y=yield,group=Block))+
      geom_bar(stat="identity",aes(fill=Variety))+
      facet_grid(Block~Variety,scales="free")+
      labs(title="The Split Plot of Oats")
```
<br><br>
Total of 18 subplots (6 levels of **Block** and 3 levels of **Variety**) are in the split plot. 

```{r}
ggplot(Oats,aes(x=Block,y=yield,group=Variety))+
      geom_line(stat="identity",aes(col=nitro))+
      facet_grid(nitro~Variety,scales="free")+
      labs(title="The Split Plot of Oats")+
      guides(col=F)
```

```{r}
mod0.1=lmer(log(yield)~nitro+Variety+nitro:Variety+(1|Block),data=Oats)
arm::display(mod0.1)
```
`mod0.1` has fixed effects of `nitro*Variety`. This model has a random effect of specified with `(1|Block)` 
and is varied by an intercept of **Block**. 

```{r}
mod1.1=lmer(log(yield)~nitro+Variety+nitro:Variety+(1|Block/Variety),data=Oats)
#arm::display(mod1.1)
summary(mod1.1)
```
`mod1.1` is in a case of nested data which is **Variety** belongs to **Block**. 

```{r}
anova(mod0.1,mod1.1,refit=F) #mod1.1
```
From ANOVA, `mod1.1` is better than `mod0.1`. 

```{r,results='asis',warning=FALSE}
print(xtable(anova(mod1.1)),type="html")
```

```{r}
ggplot(Oats,aes(x=nitro,y=log(yield),group=Variety))+
      geom_point(col="indianred2")+
      facet_grid(.~Variety,scales="free")+
      labs(title="The Relation Between log(Yield) and Nitro Depend on Variety")
```
<br><br>
In the scatter plot, it is hard to clearly see the relation between log(yield) and nitrogen level. 

```{r,warning=FALSE}
ggplot(Oats,aes(x=nitro,y=log(yield),group=Variety))+
      stat_smooth(stat="identity",fill="indianred2",method="loess")+
      facet_grid(.~Variety,scales="free")+
      labs(title="The Relation Between log(Yield) and Nitro Depend on Variety")
```
<br><br>
The observations in the data is less than 1000 so the method is `loess`. 
I wouldn't say these two variables (log(**yield**) and **nitro**) have a linear relation. 
If I have to say any relation of them, I might say they have quadratic relations.

No. I couldn't say log(yield) increases linearly with nitrogen level. Each level of **Variety** has different trends of them. 

```{r}
ggplot(Oats,aes(x=yield,y=nitro,group=Variety))+
      geom_line(aes(col=Variety))+
      facet_grid(~Block,scales="free")+
      labs(title="The Split Plot of Oats")
```
<br><br>
From the model `mod1.1`, the random effect is varied by **Block** and **Variety**:
```{r}
tapply(Oats$yield,list(Oats$Block,Oats$Variety),var)
```

```{r}
ddply(Oats,.(Block,Variety),summarise,var=var(yield))
```
`tapply` and `ddply` results are same. The observations are correlated according to the variance structure; 
correlation of random effect for the same **Variety** and **Block**. 

```{r,warning=FALSE}
tuk1 <- multcomp::glht(mod1.1,linfct=mcp(nitro="Tukey"))
summary(tuk1)
```
This shows the contrast of each nitrogen level.

```{r}
tuk.cld=cld(tuk1)
old.par=par(mai=c(1,1,1.25,1),no.readonly=TRUE) #where is alphabet?
plot(tuk.cld,col=2:5)
```
<br><br>
In the post hoc analysis with Tukey method for multiple comparisons, it shows what levels are in significant 
relationship of the differences on levels of **nitro**. In the plot the letters *b* and *c* are sharing and it 
means these are not significant. The letter, *a*, however, is significant among other letters. 
The level *Golden Rain* in **Variety** is significant then gives the greatest log(yield) after adjusting for nitrogen level.

###g.
Interpret the regression parameter for nitrogen level.
<br>
From above result (f), by each nitrogen level we know which level of **Variety** variable is significant. The level *Golden Rain* in **Variety** is statistically significant among other levels. When the nitro level is 0.2 and the variety level is golden rain, this term is significant. 

```{r,results='asis',warning=FALSE}
print(xtable(confint(mod1.1,1:3)),type="html")
```
It is the confidence limits of the standard deviations in the model.
