---
title: 'Resampling methods'
author: "Sungkyung Kang"
---

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE,cache=FALSE}
library(ISLR) #Default data set
library(MASS)
library(boot)
library(plotrix)
```

```{r}
set.seed(1001)
```

```{r}
de.logre<-glm(default~income+balance,data=Default,family='binomial')
summary(de.logre)
```
This is logistic regression for predicting *defualt* with predictors, *income* and *balance*.

####i,ii. Split sample set and fit multiple logistic regression
```{r}
train=sample(dim(Default)[1],0.5*dim(Default)[1]) #Total observation of Default data set 
de.logre1=glm(default~income+balance,data=Default,subset=train,family='binomial') 
summary(de.logre1)
```

####iii. Obtain a prediction of default status for each individual in the validation set
```{r}
#head(Default)
pred.each<-ifelse(predict(de.logre1,type="response",newdata=Default[-train,])>0.5,"Yes","No")
pred.each[1:10]
```

####iv. Compute validation set error 
```{r}
attach(Default)
mean((as.numeric(default)-predict(de.logre1,Default))[-train]^2) #61.2884
```
Validation set error: Estimated test Mean Sqaure Error (MSE) is 61.2884.

####1. sample(5000,2500)
```{r}
train1=sample(5000,2500)
de.logre2=glm(default~income+balance,data=Default,subset=train1,family='binomial')
summary(de.logre2)
pred.each<-ifelse(predict(de.logre2,type="response",newdata=Default[-train,])>0.5,"Yes","No")
pred.each[1:10]
mean((as.numeric(default)-predict(de.logre2,Default))[-train]^2)
```
Validation set error: Estimated test Mean Sqaure Error (MSE) is 64.5004.

####2. sample(1000,500)
```{r}
train2=sample(1000,500)
de.logre3=glm(default~income+balance,data=Default,subset=train2,family='binomial')
summary(de.logre3)
pred.each<-ifelse(predict(de.logre3,type="response",newdata=Default[-train,])>0.5,"Yes","No")
pred.each[1:10]
mean((as.numeric(default)-predict(de.logre3,Default))[-train]^2)
```
Validation set error: Estimated test Mean Sqaure Error (MSE) is 42.63842.

####3. sample(100,50)
```{r}
train3=sample(100,50)
de.logre4=glm(default~income+balance,data=Default,subset=train3,family='binomial')
summary(de.logre4)
pred.each<-ifelse(predict(de.logre4,type="response",newdata=Default[-train,])>0.5,"Yes","No")
pred.each[1:10]
mean((as.numeric(default)-predict(de.logre4,Default))[-train]^2)
```
Validation set error: Estimated test Mean Sqaure Error (MSE) is 761.864.

```{r}
de.logre5<-glm(default~income+balance+student,data=Default,family='binomial') #Include student
mean((as.numeric(default)-predict(de.logre5,Default))[-train]^2)
```
When the test error including *student*, MSE is 58.74683 but MSE is 61.2884 after removing dummy variable, 
*student*. So, it is measured to include dummy variable, *student* for better model with smaller MSE test 
error. 

```{r}
set.seed(3)
```

```{r}
de.logre6<-glm(default~income+balance,data=Default,family='binomial')
coef(de.logre6)
```
These are coefficients.
In this model, $default$ = $-1.154047e+01+2.080898e-05*income+5.647103e-03*balance$. 

```{r}
sum.logre6<-summary(de.logre6)
#se<-sqrt(diag(vcov(de.logre3))) 
sum.logre6$coefficients[,2] #Std.Error
```
These are standard errors of each variable.

```{r}
boot.fn=function(data,index){
  model<-glm(default~income+balance,data,family='binomial',subset=index)
return(coef(model))
} #data[,3]: balance, data[,4]: income
boot.fn(Default,default)  
```

```{r}
library(boot)
coef<-boot.fn(Default,default) 
set.seed(100)
DEF<-ifelse(Default$default=="No",0,1)
boot.fn(Default,sample(DEF,10000,replace=TRUE))
boot(data=Default,boot.fn,R=100) 
```

```{r}
sum.logre6$coefficients[,2] #Std.Error
boot(data=Default,boot.fn,R=100)
```
The first result is taken by using glm() and the last one is from bootstrap function. 
In bootstrap statistics, estimated standard errors are 'std.error'.

```{r}
data(Boston)
```

```{r}
muhat<-mean(Boston$medv) #22.53281 
```
When na.rm=TRUE or FALSE is there, there is no difference of result. An estimated for population mean 
is 22.53281.

```{r}
std=function(x)
  sd(x)/sqrt(length(x))
std(Boston$medv) #0.4088611
```
An estimate of the standard error of $\hat\mu$ is 0.4088611.

```{r}
boot.fn1=function(data,index){
  hatmu=mean(data[index])
  return(hatmu)
}
attach(Boston)
boot(medv,boot.fn1,R=1000)
```
From the bootstrap when R=1000, the standard error of 
$\hat\mu$ is 0.4148675 but result of b is 0.4088611. Both of results are quite similar. 

```{r}
se.boot1<-0.4148675 #R=1000
muhat.boot1<-22.53281 #R=1000
CI95.boot1<-c(muhat.boot1-2*se.boot1,muhat.boot1+2*se.boot1)
CI95.boot1 #21.70308:23.36255
t.test(Boston$medv) #21.72953:23.33608
```
From the results taken from bootstrap when R=1000, a 95% confidence interval 
for the mean of *medv* is [21.70308, 23.36255]. When testing t.test of *medv*, 
the 95% confidence interval is [21.72953, 23.33608]. Hence the result of bootstrap 
is wider than of t.test. 

```{r}
mean(median(Boston$medv)) #21.2
```
An estimate $\hat\mu_{med}$ for the median value of *medv* is 21.2.

```{r}
boot.fn2=function(data,index){
  hatmume=mean(median(data[index]))
  return(hatmume)
}
boot(medv,boot.fn2,R=1000)
```
When using bootstrap at R=1000, an estimated standard error of median is 0.3835767.

```{r}
mu0.1<-quantile(Boston$medv,c(.10)) #12.75 
```
The quantity of $\hat\mu_{0.1}$ is 12.75.

```{r}
boot.fn3=function(data,index){
  quan=quantile(data[index],c(.10)) #Not using $
  return(quan)
}
boot(medv,boot.fn3,R=1000)
```
The standard error of $\hat\mu_{0.1}$ is 0.5180969 when R=1000.
