---
title: 'Unsupervised learning'
author: "Sungkyung Kang"
---

```{r}
library(base)
data(USArrests)
```

```{r}
#head(USArrests)
#dim(USArrests) 50 4
pr.out.USA=prcomp(USArrests,scale=TRUE) #Recommended Scale
#summary(USArrests)
```

```{r}
biplot(pr.out.USA,scale=0)
```
<br>
This plot shows first two principal components. 

```{r}
pr.out.USA$rotation=-pr.out.USA$rotation
pr.out.USA$x=-pr.out.USA$x
biplot(pr.out.USA,scale=0)
```
<br>
This plot is the same as upper plot but a mirror image with showing 
different directions of four variables.

```{r}
pr.var.USA=pr.out.USA$sdev^2
pve.USA=pr.var.USA/sum(pr.var.USA)
pve.USA #0.62006039 0.24744129 0.08914080 0.04335752
```
With the variances of each principal components, PVE can be calculated. 
The first principal component explains 62.0% of the variance in the data, 
the second principal component explains 24.7% of the variance and so on. (Proportion of Variance Explained)

```{r}
par(mfrow=c(1,2))
plot(pve.USA,xlab="Principal component",ylab="Variance explained",ylim=c(0,1),type="o",main="PC with variance in USArrests",col="red",pch=16)
plot(cumsum(pve.USA),xlab="Principal component",ylab="Cumulative proportion of variance explained",ylim=c(0,1),type="o",pch=16,col="red",main="PC with cumulative proportion variance in USArrests")
```
<br>
The larger principal component is, the lower variance is.

```{r}
dev.off()
```

```{r}
pi.loading<-pr.out.USA$rotation
scale.USA=scale(USArrests,scale=TRUE) #Scale, center
denom.USA=sum(apply(as.matrix(scale.USA)^2,2,sum)) #col=2 
nom.USA=apply((as.matrix(scale.USA)%*%pi.loading)^2,2,sum) #(Order?)
#nom.USA1=sum((apply(as.matrix(scale.USA)%*%pi.loading,2,sum))^2) #Not working
nom.USA/denom.USA #0.62006039 0.24744129 0.08914080 0.04335752 
```
*scale()* funciton is to be centered or scale of variables matrix.
The two results (a, b) are same. 

```{r}
hc.complete=hclust(dist(USArrests,method="euclidean"),method="complete")
plot(hc.complete,main="Complete linkage of USArrests dendrogram")
```
<br>
The denderogram looks to be balanced well.

```{r}
height.3=cutree(hc.complete,3)
height.3
```
For example, *'Alabama'* instance belongs to class 1, 
*'Colorado'* instance belongs to class 2 and *Connecticut* instance belongs to class 3.

```{r}
scale.USA=scale(USArrests)
plot(hclust(dist(scale.USA,method="euclidean"),method="complete"),main="Hierarchical clustering with scale")
```
<br>
The denderogram looks to be balanced well.

```{r}
par(mfrow=c(1,2))
plot(hc.complete,main="Complete linkage (non-scale)")
plot(hclust(dist(scale.USA,method="euclidean"),method="complete"),main="Complete linkage (scale)")
```
<br>
Left panel is non-scale of complete linkage and right panel is scale of comlete linkage. 
For cutting (making clusters), sclae is better to do. <br>
With scaling, the height is adjusted.

```{r}
dev.off()
```

```{r}
table(cutree(hc.complete,3)) #non-scale
table(cutree(hclust(dist(scale.USA,method="euclidean"),method="complete"),3)) #scale
```
In this table, for the number of three clusters, each cluster with non-scale has 
similar instances rather than with scale. 

```{r}
set.seed(2)
class=c(rep(1,20),rep(2,20),rep(3,20))
ds=matrix(rnorm(20*3*50),ncol=50)
ds[class==1,]=ds[class==1,]
ds[class==2,]=ds[class==2,]-5
ds[class==3,]=ds[class==3,]+5
par(mfrow=c(1,1))
plot(ds[,1:2],col=(4-class),pch=16,xlab="",ylab="",main="3 classes")
```

```{r}
pr.out.ds=prcomp(ds)
#Not working a) to b) so have to change simulation here
plot(pr.out.ds$ds[,1:2],col=(4-class),pch=16,xlim=c(-4,4),ylim=c(-4,4),xlab="P1",ylab="P2")
```

```{r}
k3=kmeans(ds,3,nstart=30) #3 clusters, default
table(class,k3$cluster)
```
If the components showed only diagnomal like (1,1), it is perfectly done for clustering. 
In this table, observations are clustered well.

```{r}
k2=kmeans(ds,2,nstart=30) #2 clusters
table(class,k2$cluster)
```
One cluster is merged to other clusters and total classes are three.

```{r}
k4=kmeans(ds,4,nstart=30) #4 clusters
table(class,k4$cluster)
```
One more cluster is formed.

```{r} 
k3.out=kmeans(pr.out.ds$ds[,1:2],3,nstart=30) ###QN) pr.out.ds$ds[,1:2] is NULL?! #NULL #Not working
table(class,k3.out$cluster) #Not working
table(class,k3$cluster)
```
pr.out.ds$ds[,1:2] is not shown so PCA is not known as the result.

```{r}
k3scale=kmeans(scale(ds),3,nstart=30)
table(class,k3scale$cluster) #scale 
table(class,k3$cluster) #non-scale
```
In table comparison, non-scale is better than scale (shape of diagonal).

```{r}
summary(gene)  
head(gene)  
dim(gene) #1000 40
```

#####Complete method
```{r}
cd=as.dist(1-cor(gene))
hc.complete.gene=hclust((cd),method="complete")
plot(hc.complete.gene)
rect.hclust(hc.complete.gene,k=2)
```
<br>
With 40 variables, a hierarchical clustering dendrogram used method *complete* is. 
In this dendrogram, there are two distinct groups when a threshold is nearly 1.1. 
The number of groups making clusters are depended on position of threshold. 

#####Average method
```{r}
hc.average.gene=hclust((cd),method="average")
plot(hc.average.gene)
rect.hclust(hc.average.gene,k=2)
```
<br>
For making two groups, the threshold should be at height 1. However the shape of this 
dendrogram seems to be not balanced. 

#####Single method
```{r}
hc.single.gene=hclust((cd),method="single")
plot(hc.single.gene)
rect.hclust(hc.single.gene,k=2)
```
<br>
The shape of this dendrogram seems to be not balanced and not good. 
#####Ward.D method
```{r}
hc.wardD.gene=hclust((cd),method="ward.D")
plot(hc.wardD.gene)
rect.hclust(hc.single.gene,k=2)
```
<br>
The shape of this dendrogram seems to be not balanced. 

#####Single method
```{r}
hc.med.gene=hclust((cd),method="median")
plot(hc.med.gene)
rect.hclust(hc.single.gene,k=2)
```
<br>
The shape of this dendrogram seems to be not balanced and not good. 
#####Centroid method
```{r}
#hc.cent.gene=hclust((cd),method="centroid")
#plot(hc.cent.gene)
#rect.hclust(hc.centroid.gene,k=2)
```
The shape seems to be worse.

#####Among those methods, 'complete' method looks great on this dataset.

```{r}
pr.out.gene=prcomp(t(gene)) #Transpose, 100obs, 40variable
loading=apply(pr.out.gene$rotation,1,sum) #Check
label=order(abs(loading),decreasing=TRUE) #Check
head(label)
loading[label[1:20]]
```
